---
title: "Tutorial"
output: html_vignette
vignette: >
  %\VignetteIndexEntry{kNN tutorial}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Include the kNN package:

```{r}
library(kNN)
```

Load intrinsic iris dataset and store them into train, test and cl (class).

```{r}
train <- rbind(iris3[1:25,,1], iris3[1:25,,2], iris3[1:25,,3])
test <- rbind(iris3[26:50,,1], iris3[26:50,,2], iris3[26:50,,3])
cl <- factor(c(rep("s",25), rep("c",25), rep("v",25)))

```

After that, calling kNN, use kNN or kNN_R to get knn results.

```{r}
start = Sys.time()
preds1 = kNN_R(train, test, cl, k = 15, prob=TRUE, use.all = FALSE)
probs1=attributes(preds1)$prob
end = Sys.time()
print("The time cost of R-based implementation: ")
t1=end-start
print(t1)
```

```{r}
start = Sys.time()
preds2 = kNN(train, test, cl, k = 15, prob=TRUE, use.all = FALSE)
probs2=attributes(preds2)$prob
end = Sys.time()
print("The time cost of Rcpp-based implementation: ")
t2=end-start
print(t2)
```
See the acceleration rate.

```{r}
cat("The acceleration rate is", as.numeric(t1)/as.numeric(t2))

```

See the accuracy of the algorithm.

```{r}
cl_gt <- factor(c(rep("s",25), rep("c",25), rep("v",25)))
accuracy1 = sum(preds1 == cl_gt) / length(cl_gt)
accuracy2 = sum(preds2 == cl_gt) / length(cl_gt)
cat("The accuracy of R-based implementation:", accuracy1, "\n")
cat("The accuracy of Rcpp-based implementation:", accuracy2)
```

# Correctness and Effciency

## Example of iris

```{r}
# install.packages("class")
library(class)
start = Sys.time()
preds3=knn(train, test, cl, k = 15, prob=TRUE, use.all = FALSE)
probs3=attributes(preds3)$prob
end = Sys.time()
t3=end-start
print(t3)

```
### Efficiency

```{r}
cat("The acceleration rate of kNN comparing to knn in `class` package is", as.numeric(t3)/as.numeric(t2))
```

We can see our implementation accelerated by multi-thread is faster than the original knn function.

### Correctness
```{r}
sum(preds3 != preds2)
sum(preds3 != preds1)
sum(preds1 != preds2)
sum(attributes(preds3)$prob != attributes(preds2)$prob)
sum(attributes(preds3)$prob != attributes(preds1)$prob)
sum(attributes(preds1)$prob != attributes(preds2)$prob)

```

We can see all prediction labels and their corresponding probabilities are the same for each implementations.

The dataset is small and the acceleration is not remarkable. So we simulate another larger dataset.

## Example of simulated data

Create simulated data by random uniform distribution, which is a training set consisting of 2-class and 4000 samples and a test data consisting 400 samples.

```{r}
d = 2
m = d*2000
train <- matrix(runif(m*40, -1000, 1000), m, 40)
cl <- factor(c(rep(1,m/d), rep(2,m/d)))
test <- matrix(runif(m/10*40, -1000, 1000), m/10, 40)

```

Run the three implemented functions on the data

```{r}

start = Sys.time()
r1=knn(train, test, cl, k = 11, prob=TRUE, use.all = F)
end = Sys.time()
t11 = end-start
print(t11)

start = Sys.time()
r2=kNN_R(train, test, cl, k = 11, prob=TRUE, use.all = F)
end = Sys.time()
t22=end-start
print(t22)

start = Sys.time()
r3=kNN(train, test, cl, k = 11, prob=TRUE, use.all = F)
end = Sys.time()
t33=end-start
print(t33)

```

### Correctness

```{r}
sum(r1 != r2)
sum(r1 != r3)
sum(r3 != r2)
sum(attributes(r3)$prob != attributes(r1)$prob)
sum(attributes(r3)$prob != attributes(r2)$prob)
sum(attributes(r1)$prob != attributes(r2)$prob)

```

We can see in this simulated data, all predictions and probabilities are still the same.

### Efficiency
```{r}
cat("The acceleration rate of kNN comparing to knn in `class` package is", as.numeric(t11)/as.numeric(t33))

cat("The acceleration rate of kNN comparing to kNN_R is", as.numeric(t22)/as.numeric(t33))

```

We can see our implementation accelerated by multi-thread is about 2x to 3x faster than the original knn function. And we can see our implementation accelerated by multi-thread is about 100x to 300x faster than the R-based kNN_R function.


